Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-05 android-stacks.tex
square content

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptional conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by a JNI interface function;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
as the root cause of an exception stack trace.
This study provides insights on such fragilities and the robustness threats  
they impose not only to Android apps but also to other systems 
based on Java exception model.

@@

@@

Along the recent years, we have witnessed an astonishing increase in the number of
mobile applications. If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face an increasing number of threats to application robustness
 arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
 memory and battery restrictions~\cite{Zhang12}; among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions. 
Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
 own device)  reuses the embedded Java exception model to detect and handle 
 exceptional conditions.

Studies have shown, however, that the exception-related code is generally poorly understood and the
 less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
As a consequence they may affect the system the other way around.
The exception-related code may favor the introduction of failures such as 
uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying} - 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

@@

Until recent years, the availability of data, reporting such failures, was scarce. 
It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their version control systems as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Google Code.
Developers and users of open-source projects uses such provided
 infra-strucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
the software development.

@@

This study performs a post mortem analysis of the exception stack traces embedded on issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study was to: investigate whether the reported exception stack traces
 can reveal common fragilities on the exception-related code.  
We call ``fragilities" characteristics on the exception-related code 
that favor the introduction of failures such as uncaught 
exceptions -  contributing to decrease the applicaiton robustness.
To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented with additional information
extracted from JDK, Android Plaform and the applications using byte code 
and source code analysis. Manual inspection was used to leverage the understanding 
of exception stack traces and support further discussions and insights.

Some outcomes consistently detected during this study were the following:

@@

 \item  A multitude of programming mistakes were reported as the main causes of exception stack traces:
 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.

   \item  Cross-type exception wrappings (e.g., an OutOfMemoryError wrapped in a checked exception)
were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.

   \item Undocumented runtime exceptions signaled by third-party code (i.e., libraries and Android platform). 
    In this study, none of the third-party methods explicitly signaling runtime exceptions included such exceptions
    on ``exception specifications" (i.e., a list of exceptions that a method might raise during its execution).

   \item Undocumented checked exception signaled by a  JNI interface function defined on Android Platform. 
    Some flows reported a checked exception thown by a JNI interface function, and  such exception was not  
    part of the ``exception interface" of JNI function nor the methods on its invokation chain - it can lead to 
   uncaught exceptions that are difficult to debug.

@@

Such outcomes point to threats not only to the developement of robust Android apps, 
but also to the developement of any robust Java-based system. For instance, in the absence of the 
``exception specification" of  a third-party code, it is difficult or 
even enfeasible for the developer to protect the code against ``unforeseen" exceptions. 
Since, in such cases the client usually do not have access to the source code, the undocumented exceptions may 
remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
(e.g., OutOfMemoryError is wrapped in a checked Exception) 
may lead to unintended handling problems - trying to handle an instance of OutOfMemoryError ``hiden" in 
a checked exception may lead the program to an unpredictable state.

Hence, the study results are relevant to Android and 
Java developers who may underestimate how likely are the
 exceptions related to programming mistakes, and how 
 tricky it can be to prevent the uncaught exceptions from third-party code
 in the absence of ``exception interfaces". Moreover, results are relevant to 
designers of languages and tools to consider ways of reducing such fragilities 
in Android and Java environments.

The remaining of this paper is organized as follows. Section 2 presents a
background on Android Platform and Java exception model. 
Section 3 presents the study design. Section 4 reports the study findings. 
Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 concludes the paper and outlines
directions for future work.

@@

%The Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
%And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
%abnormal computation states.

The Android Application Framework supports an event-driven development model. 
Among the basic elements that comprise an Android application, we can cite the
 Activity class. The Activity is the main class of an Android app (there is usualy one
Activity per application window). Each Activity defines a set of hook methods  
(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
framework (by inversion of control), in response to infrastructure and application events.

The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent development (e.g. AsyncTask). 
Such APIs particularly useful due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsive due to resource limitation and excessive network access).

@@

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}. The checked exception received such name
 because they must be declared on the method's exception interface (i.e., a list of exceptions that a method 
might raise during its execution) and the compiler statically checks if
 appropriate handlers are provided within the system.
Both runtime exceptions and errors are also known as unchecked exceptions, because 
they do not need to be specified on method exception interface and do not trigger any 
compile time checking.

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked (because doing so 
the callers of a method will know about the exceptions that a method can throw and so 
they can decide what to do about them). However, runtime exceptions can also be 
used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate},
\footnote{152 questions in Stackoverlow are related to this debate}.

\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
  \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}

Besides, the runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references).\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. }  
The programs are not expected to handle such runtime exceptions signaled by the runtime environment ~\cite{gosling2000java}.

%In Java an exception can be thrown in one of the following
%circumstances~\cite{gosling2000java}: 
%(i) explicitly thrown when a throw statement is executed; 
%(ii) implicitly thrown by the JVM when the evaluation of an expression
% violates the normal semantics of language, also referred as a coding error
% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
%OutofMemoryError). 
%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}

%When an exception is thrown, it causes a transfer of control from the point
%where the exception occurred to a point that can be specified by the programmer
%(exception handler) - in Java it is represented by the try-catch block .

@@

\begin{figure} \centering \includegraphics[scale=0.6]{stacktrace_bw.png}
\caption{Example of an exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

%<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
%PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>

% \subsection{Public Project Repositories and Issue Trackers} 
% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
% As mentioned before, currently more and more projects are becoming available in 
% public project repositories such as GitHub and Google Code.
% Such repositories provides facilities such as....
% The issue trackers available in such projects ....

@@

\emph{II-Error represents an unrecoverable conditions which should not he handled} 
(\cite{gosling2000java}).  Errors should result from failures detected
by the runtime environemt which indicate resource deficiencies, invariant
failures or other conditions, that make impossible for the program to recover.

@@

%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
%possible to the problem}~\cite{wirfs2006toward}. 
%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
%Moreover, 
%Close to the place that the exception was throw the caller usually has
%enough contextual information to perform  corrective actions. When the exception
%is propagated far away from the method where it was signaled (i.e. further away
%#from the problem) it can be difficult to take meaningful recovery actions.
%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}

@@

\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions, specialy when developing
 library/framework reusable code. Doing so, the clients of such ibrary/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

If the developer fails to do follow such practices (specially when developing 
utility code) it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against unforeseen exceptions thrown by the utility code,
which can lead to the uncaught exceptions - one of the main causes of 
systems crashes. These best practices guided our exploratory study dscribed next.

%Next section discusses the
%study procedure and the practices' misuses that emerged from this study.

%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
%In Java when an application crashes due to an uncaught exception ... 
%Exception Stach trace is usually (procurar por artigo).

%To answer this question we selected 
%The goal of this exploratory mining study was to extract exception stack traces embedded on issues of Android projects 
%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
%Our study focused on open-source apps, since the information needed to perform our study
%cannot be retrieved from commercial apps, whose issue report systems and 
%source codes are not publicly available. The Android open-source apps were also 
%target of other research studies ~\cite{Linar13,ahimed}.

%extract exception stack traces embedded on issues of Android projects 
%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
%Our study focused on open-source apps, since the information needed to perform our study
%cannot be retrieved from commercial apps, whose issue report systems and 
%source codes are not publicly available. The Android open-source apps were also 
%target of other research studies ~\cite{Linar13,ahimed}.

@@

@@

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
of exception stack traces can pinpoint best practices violations. As mentioned before, 
call ``fragilities" characteristics on the exception-related code that favor the introduction
of failures such as uncaught exceptions and unintended handling.

To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
which extracts the exception stack traces embeeded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code.
Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

@@

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories. Running the ExceptionMiner tool (described next)
 in this set we observed that 589 apps had at least one issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
on the issues including exception stack traces.

% \centering
%  \begin{tabular}{|p{2cm}| p{5cm}|}
%    \hline

\begin{table}
  \centering
  \begin{tabular}{lr|lr}
    \hline
     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
    \hline
empty &	54,24\% & Defect &	91,96\% \\
Defect &	39,56\%  & Enhancement  &	3,16\% \\
Enhancement &	0,57\% & Task	& 1,37\% \\
Support &	0,52\% & empty &	1,12\% \\
Problem &	0,36\% & StackTrace &	0,70\% \\
Others &	4,74\% &  Others &	1,68\% \\   
  \hline
  \end{tabular}
  \caption{Lables on issues including exception stack traces.}
  \label{tab:lables}
\end{table}

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
issues from our analysis.

>>>>

\begin{abstract}
%The number of mobile apps is increasing in a daily rate. If on the one hand they
%extend phones capabilities far beyond of the basic calls, on the other hand they have to 
%cope with an increasing number of exceptional conditions 
%(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
%Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
%fundamental part of such apps. However, studies have shown that such mechanisms are the less 
%understood and tested parts of the system. As a consequence they may affect the system the 
%other way around, making the system even less robust, leading to failures such as uncaught exceptions 
%and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).

This paper reports a study which mined the Java stack traces included in
159,048 issues reported on Android projects hosted in GitHub (482 projects) 
and Google Code (157 projsects). 
The goal of this study is to investigate  whether the exception handling logic on these stack traces
 can reveal common characteristics on the exception-related code that
are particularly fault prone and may lead to a decrease in application robustness.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Key problems identified in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by a JNI interface function;
(3) unexpected exception wrappings (such as Errors being wrapped in checked exceptions),
indicating that Java's hybrid exception model is not fully used in accordance with its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
as the root cause of an exception stack trace, suggesting the presence of programming mistakes.
This study provides insights on such fragilities and the robustness threats  
they impose not only to Android apps but also to other systems 
based on Java exception model.

@@

%Total page numbers then:
%intro:       1
%background:  1.5
%method:      1
%tool         0.5
%findings     4
%discussion   1
%threats      .5
%conc         .5
%------------------+
%total        10 pages

@@

In recent years, we have witnessed an astonishing increase in the number of
mobile applications. These applications extend phones capabilities 
far beyond basic calls and textual messages. Such applications, however,
have to face an increasing number of threats to application robustness
 arising from: failures in underlying middleware and hardware (e.g., camera, sensors);
failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
 memory and battery restrictions; and noisy external resources ~\cite{Zhang12} (e.g., wireless connections,GPS, bluetooth).

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions.
In this paper we will be concerned with exception handling in Android apps,
which reuse Java's exception handling model.

%Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
%model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
% own device)  reuses the embedded Java exception model to detect and handle 
% exceptional conditions.

Studies have shown that exception-related code is generally poorly understood and among the least tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
As a consequence they may inadvertently negatively affect the system: Exception-related code may introduce failures such as 
uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

@@

%% I am not so sure this data was not available. Let's omit this paragraph
%Until recently, the availability of data reporting such failures was scarce. 
%It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
%Recently we have observed a rapid increase in the number of open
%source projects that make their version control systems as well as the issue tracking systems
%publicly available in repository hosting sites such as GitHub and Google Code.
%Developers and users of open-source projects uses such provided
% infrastrucute to share information (e.g., exception stack %traces~\cite{bettenburg2008makes,schroter2010stack}) across 
%the software development.

@@

This study performs a post mortem analysis of the exception stack traces 
included in issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common fragilities in the exception-related code.  
We call ``fragilities'' characteristics of the exception-related code 
that increase the likelihood of introducing failures such as uncaught 
exceptions -  leading to a decrease in application robustness.

To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of byte code 
and source code analysis
 with additional exception-related information
extracted from the Android platform and Android applications.

Outcomes of our study include:

%Manual inspection was used to leverage the understanding 
%of exception stack traces and support further discussions and insights.

% Some outcomes consistently detected during this study were the following:

@@

       \item Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause.

       \item  4.4\% of the reported exception stack traces are ``undocumented runtime exceptions'' raised by third-party code. 
             (35 methods from the Android platform and 44 from libraries).

     \item  We detected several cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
in stack traces. This suggests that the purpose of Java's hybrid exception model comprising checked and unchecked exceptions may not have been adequately used.

   \item Undocumented checked exception signaled by a  JNI interface function defined on Android platform. 
    Some flows reported a checked exception thrown by a JNI interface function, yet this exception was not  
    part of the ``exception interface'' of the JNI function nor the methods on its invocation chain. This can lead to 
   uncaught exceptions that are difficult to debug.

@@

Such outcomes point to threats not only to the development of robust Android apps, 
but also to the development of any robust Java-based system. For instance, in the absence of the 
``exception specification" of third-party code, it is difficult or 
even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
Since in such cases the client usually does not have access to the source code, such undocumented exceptions may 
remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
(e.g., OutOfMemoryError is wrapped in a checked Exception) 
may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hidden" in 
a checked exception may bring the program to an unpredictable state.

Hence, the study results are relevant to Android and 
Java developers who may underestimate how 
 exceptions are related to programming mistakes, and how 
 tricky it can be to prevent the uncaught exceptions from third-party code
 in the absence of ``exception interfaces". Moreover, the results are relevant to 
designers of languages and tools to consider ways of reducing such fragilities 
in Android and Java environments.

The remaining of this paper is organized as follows. 
Section~\ref{sec:back} presents a
background on Android platform and Java exception model. 
Section~\ref{sec:study} presents the study design. 
Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
Section~\ref{sec:result} reports the findings of our study.
Section~\ref{sec:disc} provides a discussion of the wider implications of our results.
Section~\ref{sec:threats} presents the threats to validity associated to this study. 
Finally Section~\ref{sec:rele} describes related work, 
and Section~\ref{sec:conc} concludes the paper and outlines
directions for future work.

@@

The Android Application Framework supports an event-driven development model. 
Among the basic elements that comprise an Android application, we can cite the
 Activity class. The Activity is the main class of an Android app (there is usually one
Activity per application window). Each Activity defines a set of hook methods  
(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
framework (by inversion of control), in response to infrastructure and application events.
The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent development (e.g. AsyncTask) - 
which are usually needed due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsive due to resource limitation and excessive network access).

@@

In Java, exceptions are represented according to a class hierarchy, on which
 every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception), the runtime exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}. The checked exception received such name
 because they must be declared on the method's exception interface (i.e., a list of exceptions that a method 
might raise during its execution) and the compiler statically checks if
 appropriate handlers are provided within the system.
Both runtime exceptions and errors are also known as unchecked exceptions, because 
they do not need to be specified on method exception interface and do not trigger any 
compile time checking.

By convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The user-defined exceptions 
can be either checked or runtime. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked, because doing so 
the callers of a method will know about the exceptions that a method can throw, and then 
they can decide what to do about them. However, runtime exceptions can also be 
used to represent user-defined exceptional conditions.There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverlow are related to this debate}.

The runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references).\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. }  
The programs are not expected to handle such runtime exceptions signaled by the runtime environment ~\cite{gosling2000java}.

%\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
% \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}

@@

\begin{figure} \centering \includegraphics[scale=0.5]{stacktrace_bw.png}
\caption{Example of an exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

@@

\emph{II-Error represents an unrecoverable conditions which should not be handled} 
(\cite{gosling2000java}).  Errors should result from failures detected
by the runtime environment which indicate resource deficiencies, invariant
failures or other conditions, that make impossible for the program to recover.

@@

@@

\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
The exceptions thrown by a method are an important part of methods interface,
and is required to use the method properly. The checked exceptions are already
part of the  methods signature, and the method caller is aware of the checked
exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
also wise to document  (either or on the exception interface or using @throws tag on Javadoc) 
the explicitly thrown runtime exceptions\footnote{This
excludes the implicitly signaled runtime exceptions related to programming
mistakes} as carefully as checked exceptions, e when developing
 library/framework reusable code. Doing so, the clients of such library/framework
will be aware of all the exceptions a method can throw and can design the code
to deal with them ~\cite{Robil00}.

If the developer fails to follow such practices (specially when developing 
library/framework code) it will be difficult or even impossible for the caller to 
make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
and protect the system against unforeseen exceptions thrown by the utility code,
which can lead to the uncaught exceptions - one of the main causes of 
systems crashes. These best practices guided our exploratory study described next.

@@

\begin{figure*} \centering \includegraphics[scale=0.3]{studyoverview3.png}
\caption{Study overview.} \label{fig:overview} \end{figure*}

@@

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{common fragilities} on the exception code of both the applications and the underlying  framework. 
Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
of exception stack traces and the related source code it can pinpoint best practices violations. As mentioned before, 
we call  ``fragilities" characteristics on the exception-related code that favor the introduction
of failures such as uncaught exceptions and unintended handling.

To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

@@

\begin{table}
  \centering
  \begin{tabular}{lr|lr}
    \hline
     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
    \hline
empty &	54.24\% & Defect &	91.96\% \\
Defect &	39.56\%  & Enhancement  &	3.16\% \\
Enhancement &	0.57\% & Task	& 1.37\% \\
Support &	0.52\% & empty &	1.12\% \\
Problem &	0.36\% & StackTrace &	0.70\% \\
Others &	4.74\% &  Others &	1.68\% \\   
  \hline
  \end{tabular}
  \caption{Labels on issues including exception stack traces.}
  \label{tab:lables}
\end{table}

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
an off-line mirror of the data  offered through the Github API.  
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  
Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
 find 2.542 repositories. Running the ExceptionMiner tool 
 in this set we observed that 589 projects had at least one issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces.

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
issues from our analysis.

\subsubsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
 however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on Google Code search interface) for the term ``android". On January 2014, when we queried Google Code, such heuristic could
 returned a list of 788  projects. This list comprised the seeds sent to our Crawler .
